SULI Report!

### 2015-08-18 doc/suli-report-swe.tex
Added a first draft of the Results section to the SULI report. #81

\begin{abstract}
As dark matter does not absorb or emit light, its distribution in the universe must be inferred through indirect effects such as the gravitational lensing of distant galaxies. While most sources are only weakly lensed, the systematic alignment of background galaxies around a foreground lens can constrain the mass of the lens which is largely in the form of dark matter. In this paper, a simple halo model is used to approximate the dark matter mass distribution of the Millennium Simulation on the galaxy and cluster scales. Background sources are generated at a redshift of ${z = 1.3857}$ and then weakly lensed by the foreground halos contained in a light cone centered along each line of sight, which is carried out by the \texttt{Pangloss} framework\footnote{https://github.com/drphilmarshall/Pangloss}. The model-predicted ellipticities of the background galaxies after lensing are compared to the ``true'' weakly-lensed ellipticities determined by the ray-traced convergence and shear calculated by Hilbert et al. (citation). Using the shear-shear and galaxy-galaxy correlation functions, I find that the \texttt{Pangloss} framework systematically under-predicts the shear power in both statistics and does not accurately capture the effect of dark matter structure at scales larger than 1 arcminute. Physical and computational shortcomings of the \texttt{Pangloss} framework are discussed, as well as potential improvements for upcoming work.
\end{abstract}

@@

Instead of comparing the ray-traced and \texttt{Pangloss}-predicted lensed ellipticities for individual galaxies, the lensing is characterized globally with correlation functions. The ellipticity-ellipticity correlation function measures how correlated the ellipticities of pairs of galaxies are as a function of separation distance, while the galaxy-galaxy correlation function measures the correlation of lensed ellipticities around foreground halos as a function of separation distance. For readers that are unfamiliar with correlation functions in the context of cosmology or want a visual aid, see Appendix A. Both correlation functions are used in this work to measure how well the \texttt{Pangloss} framework models weak lensing by dark matter structures using the publically available \texttt{TreeCorr} module written by Mike Jarvis\footnote{https://github.com/rmjarvis/TreeCorr}. Note that the correlation function definition used in \texttt{TreeCorr} is slightly different than that used in most of the literature; for a derivation of the connection between Jarvis's definition and the more common Schneider definition (cite), see Appendix B.

@@

\begin{itemize}
\item Maps??
\item Correlation Functions
    \begin{itemize}
    \item shear-shear
    \item galaxy-galaxy
    \item progression of each as $R$ increases??
    \item cf`s without shape noise and with shape noise
    \item cf`s with all galaxies vs with only `relevant' galaxies?
    \end{itemize}
\item The Need for Speed
    \begin{itemize}
    \item Table or graph of speedups and CPU time per lightcone progression (make sure to check \texttt{timeit} first!)
    \end{itemize}
\end{itemize}

@@

\begin{itemize}
\item P.J. Marshall Thesis?
\item Collett et al.
\item S. Hilbert (ray-traced) http://arxiv.org/abs/0809.5035
\item Millennium Simulation
\item Schneider
\item Jarvis \texttt{TreeCorr}
\item lensing calculation http://arxiv.org/pdf/astro-ph/9908213v1.pdf
\item NFW profile
\item BMO truncation, BMO vs NFW http://arxiv.org/pdf/1301.1684v2.pdf
\item SMHR http://arxiv.org/pdf/1401.7329v1.pdf
\item Planck
\item galaxy curves
\item BAOs
\item LCDM
\item GR
\item wl distortion percent
\end{itemize}

@@

\begin{itemize}

\end{itemize}

@@

>>>>

\begin{abstract}
As dark matter does not absorb or emit light, its distribution in the universe must be inferred through indirect effects such as the gravitational lensing of distant galaxies. While most sources are only weakly lensed, the systematic alignment of background galaxies around a foreground lens can constrain the mass of the lens which is largely in the form of dark matter. In this paper, a simple halo model is used to approximate the dark matter mass distribution of the Millennium Simulation on the galaxy and cluster scales. Background sources are generated at a redshift of ${z = 1.3857}$ and then weakly lensed by the foreground halos contained in a light cone centered along each line of sight, which is carried out by the \texttt{Pangloss} framework\footnote{https://github.com/drphilmarshall/Pangloss}. The model-predicted ellipticities of the background galaxies after lensing are compared to the ``true'' weakly-lensed ellipticities determined by the ray-traced convergence and shear calculated by Hilbert et al. (citation). Using the ellipticity-ellipticity and galaxy-galaxy correlation functions, I find that the \texttt{Pangloss} framework systematically under-predicts the shear power in both statistics and does not accurately capture the effect of dark matter structure at scales larger than 1 arcminute. Physical and computational shortcomings of the \texttt{Pangloss} framework are discussed, as well as potential improvements for upcoming work.
\end{abstract}

@@

Instead of comparing the ray-traced and \texttt{Pangloss}-predicted lensed ellipticities for individual galaxies, the lensing is characterized globally with correlation functions. The ellipticity-ellipticity correlation function measures how correlated the ellipticities of pairs of galaxies are as a function of separation distance, while the galaxy-galaxy correlation function measures the correlation of lensed ellipticities around foreground halos as a function of separation distance (cite corrFuns). For readers that are unfamiliar with correlation functions in the context of cosmology or want a visual aid, see Appendix A. Both correlation functions are used in this work to measure how well the \texttt{Pangloss} framework models weak lensing by dark matter structures using the publically available \texttt{TreeCorr} module written by Mike Jarvis\footnote{https://github.com/rmjarvis/TreeCorr}. Note that the correlation function definition used in \texttt{TreeCorr} is slightly different than that used in most of the literature; for a derivation of the connection between Jarvis's definition and the more common Schneider definition (cite), see Appendix B.

@@

Using the presented methodology, (x \# of background sources) background galaxies in a (y*y square deg) subset of the (0,0,0,0) Millennium Simulation foreground catalog (cite where to get data) were generated and lensed by both the ray-traced shear and convergence as well as the \texttt{Pangloss} framework. Both sets of lensed ellipticities were characterized with correlation functions, as well as the intrinsic ellipticities of the sources before lensing.

\subsection*{Ellipticity-Ellipticity Correlation Function}

The first test of the halo model was with the ellipticity-ellipticity correlation function with a lightcone radius of 2 arcminutes which is given in Figure (??). The statistic measured the average correlation between pairs of ellipticities at separation distances between (0.1??) arcminutes and 2 arcminutes. The correlation function of the intrinsic ellipticities plotted in blue shows no significant deviation from zero as expected, as before lensing the galaxies had random orientations. Both the ray-traced, plotted in green, and \texttt{Pangloss}, plotted in purple, ellipticities show positive correlation on smaller scales up to a separation of 0.8 arcminutes. However, the ray-traced correlation is positive on all calculated scales while the halo model does not predict any significant corrlelation at separations above 0.8 arcminutes. Additionally, while the shapes of the correlation functions for both sets of lensed ellipticities aggree well until 0.8 arcminutes, there is a clear systematic underprediction of shear power by the halo model on all scales. On average, the shear power is underpredicted by (??) percent.

%************************
(e-e Correlation Function Plot)
%************************

Using the same catalog of background sources, this statistic was recalculated for the ray-traced and halo model lensed ellipticities for various lightcone radii ranging from $R=1$ arcminute to $R=???$ arcminutes. The result is shown in Figure (??).

%************************
(e-e Correlation Function Plot with multiple R's)
%************************

(Analysis of plot)

\subsection*{Galaxy-Galaxy Correlation Function}

The catalog of lensed ellipticities was also analyzed using the galaxy-galaxy correlation function with separation distances from (0.1??) arcminutes to 2 arcminutes and a lightcone radius of 2 arcminutes. The result is shown in Figure (??). As with the ellipticity-ellipticity correlation function, there is no significant correlation for the intrinsic ellipticities at any scale as expected. The lensed ellipticities from the ray-tracing and halo model both have positive correlation on all calculated scales and are in excellent agreement on scales from (??) arcminutes to (??) arcminutes. However, the halo model is still systematically underpredicting the shear power on scales above (??) arcminutes, although now only by an average of (??) percent. A new feature is the large overprediction of shear power on separation scales smaller than (??) arcminutes. However, background sources this close to foreground halos are often in the strong lensing regime which is not currently accounted for by the \texttt{Pangloss} framework.

%************************
(n-g Correlation Function Plot)\\
%************************
(Make a plot of n-g correlation function with multiple R's??)

\subsection*{Computational Performance}

A secondary goal of this research was to investigate how quickly the prediction for the shear and convergence for a single background source could be made. The average CPU time required for the prediction per lightcone was (?? $\pm$ ??) for the first run. However, many small improvements to the coding infastructure of \texttt{Pangloss} were made to speed up the prediction. The average CPU time for a lensing prediction per lightcone after various changes is plotted in Figure (??).

%************************
(CPU time per lightcone plot)\\
%************************
(Include RAM in plot??)

@@

\begin{itemize}
\item P.J. Marshall Thesis?
\item Collett et al.
\item S. Hilbert (ray-traced) http://arxiv.org/abs/0809.5035
\item Millennium Simulation
\item Schneider
\item Jarvis \texttt{TreeCorr}
\item lensing calculation http://arxiv.org/pdf/astro-ph/9908213v1.pdf
\item NFW profile
\item BMO truncation, BMO vs NFW http://arxiv.org/pdf/1301.1684v2.pdf
\item SMHR http://arxiv.org/pdf/1401.7329v1.pdf
\item Planck
\item galaxy curves
\item BAOs
\item LCDM
\item GR
\item corrFuns in cosmology
\item wl distortion percent
\end{itemize}

@@

@@

%% Results
\begin{itemize}
\item Maps??
\item Correlation Functions
    \begin{itemize}
    \item shear-shear
    \item galaxy-galaxy
    \item progression of each as $R$ increases??
    \item cf`s without shape noise and with shape noise
    \item cf`s with all galaxies vs with only `relevant' galaxies?
    \end{itemize}
\item The Need for Speed
    \begin{itemize}
    \item Table or graph of speedups and CPU time per lightcone progression (make sure to check \texttt{timeit} first!)
    \end{itemize}
\end{itemize}
